### 2018-04-22 MiP3.Rmd
Adding MiP3

---
title: "MPI3"
author: "Mark Hand"
date: "12/12/2017"
output: html_document
---

```{r setup, include=F}
knitr::opts_chunk$set(echo = TRUE) # Something about loading stuff more quickly

setwd("~/Dropbox/PhD_data/MiP")
library(tidyverse) # For data manipulation and visualization
library(stargazer) # For output tables 
library(plm) # For linear panel regression
library(car) # For logging the dependent variable; might be in tidyverse already 
library(gridExtra) # For multi-plot outputs 
library(maps) # Map data. May also be in tidyverse? 
library(mapdata) # More map data? 
library(R.utils) # capitalizing? 
library(MASS) # Boxcox? May also be in tidyverse? 
```

```{r Data import, include=F}
#### Data ####
# Data from Michael Barber's work on political polarization: 
# https://doi.org/10.1086/683453
Barber <- read_csv("~/Dropbox/PhD_data/MiP/data/Barber/Barber Limits Data.csv")
# Getting 2008 only data (reasons for which will become clear in the next sections)
Barber2008 <- Barber %>% filter(year==2008)

# Data from ICC greasethewheels.org/cci/
# Downloaded as .csv, imported as is. 
convictions <- read_csv("~/Dropbox/PhD_data/MiP/data/ICS/CCI19762015.csv")
perceptions <- read_csv("~/Dropbox/PhD_data/MiP/data/ICS/CPS20142016.csv")
#Doesn't havetime series data 
reflections <- read_csv("~/Dropbox/PhD_data/MiP/data/ICS/CRI19772013.csv")
# Though I don't like the name reflections--it implies an unproven causal relationship.

#### Plotting Barber #### 
ind_limits_2010.plot <- ggplot(Barber, aes(ind_limits_2010)) + geom_density() + theme_minimal()
ind2pty_limits_2010.plot <- ggplot(Barber, aes(ind2pty_limits_2010)) + geom_density() + theme_minimal()
pac_limits_2010.plot <- ggplot(Barber, aes(pac_limits_2010)) + geom_density() + theme_minimal()
Barber.plots <- grid.arrange(ind_limits_2010.plot, ind2pty_limits_2010.plot, pac_limits_2010.plot, nrow=1)
# ![](Barber.plots.png)
```

```{r wrangling, include=FALSE}
#### Wrangling Barber and ICS ####
corruption <- full_join(convictions, reflections)
corruption <- corruption %>% 
  rename("st"="State", "year"="Year")
corruption$st <- state.abb[match(corruption$st,state.name)]
BarberICC <- left_join(Barber, corruption)

BarberICC <- BarberICC %>% 
  mutate(
    ind_limits_dummy = ifelse(ind_limits_2010==99999,0,1),
    ind2pty_limits_dummy = ifelse(ind2pty_limits_2010==99999,0,1),
    pac_limits_dummy = ifelse(pac_limits_2010==99999,0,1),
    total_limits_dummy = ifelse(total_limits_2010==99999,0,1)
  )

#### Plotting ICS ####
CRIdensity <- ggplot(BarberICC,aes(CRI)) + geom_density() + theme_minimal()
CCIdensity <- ggplot(BarberICC,aes(CCI)) + geom_density() + theme_minimal()
grid.arrange(CCIdensity, CRIdensity, nrow=1)
```

```{r ICS density plots, include=F=F}
grid.arrange(CCIdensity, CRIdensity, nrow=1)
```

```{r corruption plots, include=F}
CCI1<-ggplot(BarberICC,aes(ind_limits_2010,CCI)) + geom_point() + theme_minimal()
CCI2<-ggplot(BarberICC,aes(ind2pty_limits_2010,CCI)) + geom_point() + theme_minimal()
CCI3<-ggplot(BarberICC,aes(pac_limits_2010,CCI)) + geom_point() + theme_minimal()
CRI1<-ggplot(BarberICC,aes(ind_limits_2010,CRI)) + geom_point() + theme_minimal()
CRI2<-ggplot(BarberICC,aes(ind2pty_limits_2010,CRI)) + geom_point() + theme_minimal()
CRI3<-ggplot(BarberICC,aes(pac_limits_2010,CRI)) + geom_point() + theme_minimal()
CCIandCRI <- grid.arrange(CCI1, CCI2, CCI3, CRI1,CRI2,CRI3, nrow=2)
```

```{r govtrustgrid, include=F}
govtrustgrid
```

```{r models, include=F}
## Plotting 
hist(BarberICC$CCI)
hist(BarberICC$CRI)
hist(BarberICC$ind_limits_2010[BarberICC$ind_limits_2010<99999])

## Modeling

## CCI 
# Individual limits and corruption
CCIreg1  <- plm(CCI ~ ind_limits_2010*ind_limits_dummy,data = BarberICC, index = c("st","year"), effect="twoways")
summary(CCIreg1)
# No signifigance

# Individual-to-party limits and corruption
CCIreg2  <- plm(CCI ~ ind2pty_limits_2010*ind2pty_limits_dummy,data = BarberICC, index = c("st","year"), effect="twoways")
summary(CCIreg2)
# No Signifigance

# PAC Limits and corruption
CCIreg3  <- plm(CCI ~ pac_limits_2010*pac_limits_dummy,data = BarberICC, index = c("st","year"), effect="twoways")
summary(CCIreg3)
# No Signifigance

# Total Limits and corruption
CCIreg4  <- plm(CCI ~ total_limits_2010*total_limits_dummy,data = BarberICC, index = c("st","year"), effect="twoways")
summary(CCIreg4)
# One star. Hmm...

# Everybody? 
CCIreg5 <- plm(CCI ~ ind_limits_2010*ind_limits_dummy + ind2pty_limits_2010*ind2pty_limits_dummy + pac_limits_2010*pac_limits_dummy, data = BarberICC, index = c("st","year"), effect="twoways")

## CRI
CRIreg1  <- plm(CRI ~ ind_limits_2010*ind_limits_dummy,data = BarberICC, index = c("st","year"), effect="twoways")
summary(CRIreg1)
# No signifigance

# Individual-to-party limits and reflection
CRIreg2  <- plm(CRI ~ ind2pty_limits_2010*ind2pty_limits_dummy,data = BarberICC, index = c("st","year"), effect="twoways")
summary(CRIreg2)
# One star. Hmm...

# PAC Limits and reflection
CRIreg3  <- plm(CRI ~ pac_limits_2010*pac_limits_dummy,data = BarberICC, index = c("st","year"), effect="twoways")
summary(CRIreg3)
# No Signifigance

# All Limits and reflection
CRIreg4  <- plm(CRI ~ total_limits_2010*total_limits_dummy,data = BarberICC, index = c("st","year"), effect="twoways")
summary(CRIreg4)
# No Signifigance

# Everybody? 
CRIreg5 <- plm(CRI ~ ind_limits_2010*ind_limits_dummy + ind2pty_limits_2010*ind2pty_limits_dummy + pac_limits_2010*pac_limits_dummy, data = BarberICC, index = c("st","year"), effect="twoways")

## CRI and CCI? 
CRIonCCI <- plm(CRI ~ CCI,data = BarberICC, index = c("st","year"), effect="twoways")
summary(CRIonCCI)
CCIonCRI <- plm(CCI ~ CRI,data = BarberICC, index = c("st","year"), effect="twoways")
summary(CCIonCRI)
cor(BarberICC$CCI, BarberICC$CCI, method="pearson") # Why didn't this work? 
ggplot(BarberICC, aes(CCI,CRI)) + geom_point(alpha=.1) + geom_smooth(method="lm", se=T)
# How does one interpret these??

```

```{r stargazer, include=F}
stargazer(CCIreg1, CCIreg2, CCIreg3, CCIreg5,
  type="text", 
  title = "Are campaign limits and corruption convictions correlated?",
  out="CCI.htm",
  dep.var.labels = c("Corruption Convictions Index"),
  covariate.labels=c("Individual limit amounts", "Individual limits (dummy)","Individual-to-party limit amount","Individual-to-party limits (dummy)","PAC limit amounts","PAC limits (dummy)",
  notes = "I make this look good!")
          )

stargazer(CRIreg1, CRIreg2, CRIreg3, CRIreg5,
  type="text", 
  title = "Are campaign limits and corruption reflections correlated?",
  out="CRI.htm",
  dep.var.labels = c("Corruption Reflections Index"),
  covariate.labels=c("Individual limit amounts", "Individual limits (dummy)","Individual-to-party limit amount","Individual-to-party limits (dummy)","PAC limit amounts","PAC limits (dummy)",
  notes = "I make this look good!")
          )
```

```{r Sandbox, include=F}
BarberICC %>% 
  mutate(
        CRI=replace(CRI, which(CRI==0), NA),
        CCI=replace(CCI, which(CCI==0), NA)
  ) -> test

CCIreg1.2  <- plm(CCI ~ ind_limits_2010*ind_limits_dummy,data = test, index = c("st","year"), effect="twoways")
summary(CCIreg1.2)
#

# Individual-to-party limits and reflection
CCIreg2.2  <- plm(CCI ~ ind2pty_limits_2010*ind2pty_limits_dummy,data = test, index= c("st","year"), effect="twoways")
summary(CCIreg2.2)
#

# PAC Limits and reflection
CCIreg3.2  <- plm(CCI ~ pac_limits_2010*pac_limits_dummy,data = test, index = c("st","year"), effect="twoways")
summary(CCIreg3.2)
#

# Double checking those dummy variables # 
summary(plm(CCI ~ ind_limits_dummy + ind2pty_limits_dummy + pac_limits_dummy, data = BarberICC, index = c("st","year"), effect="twoways"))
summary(plm(CRI ~ ind_limits_dummy + ind2pty_limits_dummy + pac_limits_dummy, data = BarberICC, index = c("st","year"), effect="twoways"))
BarberICC %>% filter(ind_limits_dummy==1) -> Dummytest
summary(plm(CCI ~ ind_limits_2010*ind_limits_dummy, data = BarberICC, index = c("st","year"), effect="twoways"))
summary(plm(CCI ~ ind_limits_2010, data = Dummytest, index = c("st","year"), effect="twoways"))
# Why in the world would these two be different?

# Did I do logs? 
summary(plm(CCI ~ log(ind_limits_2010)*ind_limits_dummy + log(ind2pty_limits_2010)*ind2pty_limits_dummy + log(pac_limits_2010)*pac_limits_dummy, data = BarberICC, index = c("st","year"), effect="twoways"))
summary(plm(CRI ~ ind_limits_2010*ind_limits_dummy + log(ind2pty_limits_2010)*ind2pty_limits_dummy + pac_limits_2010*pac_limits_dummy, data = BarberICC, index = c("st","year"), effect="twoways"))
# No change in stars

# Boxcox, maybe? 
summary(boxcox(plm(CRI ~ ind_limits_2010*ind_limits_dummy + ind2pty_limits_2010*ind2pty_limits_dummy + pac_limits_2010*pac_limits_dummy, data = BarberICC, index = c("st","year"), effect="twoways"),lambda=seq(0,1,by=.1)))
summary(boxcox(plm(CCI ~ ind_limits_2010, data = Dummytest, index = c("st","year"), effect="twoways")))
# Umm #

summary(reflections$CRI)
reflections %>% 
  summarize(
    mean=mean(CRI),
    sd=sd(CRI)) 
ggplot(data=subset(reflections,State=="South Dakota"), aes(CRI)) + geom_histogram()
ggplot(data=subset(reflections,State=="New York"), aes(CRI)) + geom_histogram()
ggplot(data=subset(reflections,State=="South Dakota"), aes(Year, CRI)) + geom_line()
ggplot(data=reflections, aes(CRI)) + geom_histogram()
ggplot(reflections, aes(Year, CRI, group=Year)) + geom_boxplot()
```