### 2018-05-21 MiP4.Rmd
Adding data, sources, models

---
title: "MiP4"
author: "Mark Hand"
date: "5/21/2018"
output: html_document
---

```{r setup, include=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE, message=FALSE, warning=FALSE)
library(haven) # import .dta
library(tidyverse) # everything 
library(plm) # For linear panel regression
```

```{r data import, include=FALSE}
# Alt and Lassen's "Enforcement and Public Corruption: Evidence from US States"
# http://epstein.wustl.edu/research/LPE.Alt.pdf
AltLassen <- read_dta("data/Replication/AltLassenJLEO.dta")

# Michael Barber's "Ideological Donors, Contribution Limits, and the Polarization of American Legislatures"  
# https://doi.org/10.1086/683453
Barber <- read_csv("~/Documents/GitHub/Money-in-Politics/data/Barber/Barber_Limits_Data.csv")

# Data from ICC 
# www.greasethewheels.org/cci/
convictions <- read_csv("~/Documents/GitHub/Money-in-Politics/data/ICS/CCI19762015.csv")
reflections <- read_csv("~/Documents/GitHub/Money-in-Politics/data/ICS/CRI19772013.csv")
```

```{r data wrangling}
# Combining Barber and ICS 
corruption <- full_join(convictions, reflections)
corruption <- corruption %>% 
  rename("st"="State", "year"="Year")
corruption$st <- state.abb[match(corruption$st,state.name)]
BarberICC <- left_join(Barber, corruption)

BarberICC <- BarberICC %>% 
  mutate(
    ind_limits_dummy = ifelse(ind_limits_2010==99999,0,1),
    ind2pty_limits_dummy = ifelse(ind2pty_limits_2010==99999,0,1),
    pac_limits_dummy = ifelse(pac_limits_2010==99999,0,1),
    total_limits_dummy = ifelse(total_limits_2010==99999,0,1)
  )

# Adding AltLassen
table(AltLassen$state_name)
AltLassen$state_name[AltLassen$state_name=="Californ"] = "California"
AltLassen$state_name[AltLassen$state_name=="Connecti"] = "Connecticut"
AltLassen$state_name[AltLassen$state_name=="Louisian"] = "Louisiana"
AltLassen$state_name[AltLassen$state_name=="Massachu"] = "Massachusetts"
AltLassen$state_name[AltLassen$state_name=="Minnesot"] = "Minnesota"
AltLassen$state_name[AltLassen$state_name=="Mississi"] = "Mississippi"
AltLassen$state_name[AltLassen$state_name=="NewHamps"] = "New Hampshire"
AltLassen$state_name[AltLassen$state_name=="NewJerse"] = "New Jersey"
AltLassen$state_name[AltLassen$state_name=="NewMexic"] = "New Mexico"
AltLassen$state_name[AltLassen$state_name=="NewYork"] = "New York"
AltLassen$state_name[AltLassen$state_name=="NorthCar"] = "North Carolina"
AltLassen$state_name[AltLassen$state_name=="NorthDak"] = "North Dakota"
AltLassen$state_name[AltLassen$state_name=="Pennsylv"] = "Pennsylvania"
AltLassen$state_name[AltLassen$state_name=="RhodeIsl"] = "Rhode Island"
AltLassen$state_name[AltLassen$state_name=="SouthCar"] = "South Carolina"
AltLassen$state_name[AltLassen$state_name=="SouthDak"] = "South Dakota"
AltLassen$state_name[AltLassen$state_name=="Tennesse"] = "Tennessee"
AltLassen$state_name[AltLassen$state_name=="Washingt"] = "Washington"
AltLassen$state_name[AltLassen$state_name=="WestVirg"] = "West Virginia"
AltLassen$state_name[AltLassen$state_name=="Wisconsi"] = "Wisconsin"
AltLassen$state_name <- state.abb[match(AltLassen$state_name,state.name)]
table(AltLassen$state_name)
AltLassen <- AltLassen %>% 
  rename("st"="state_name")
AltLassen$year <-as.integer(AltLassen$year)

AltBarber <- inner_join(BarberICC, AltLassen)
```

```{r modeling, echo=FALSE}
#Next: Panel model of corruption against unemployment and high school? 
testmodel01 <- plm(CCI ~ rpcinc_annual_new + unemp + high_school + logNumpop + citi6006 + grelimp, data = AltBarber, index = c("st","year"), effect="twoways")
summary(testmodel01)

testmodel02 <- plm(CCI ~ jdiv + bindingonetermlimit + bindingtwotermlimit + rpcinc_annual_new + eousa + unemp + citi6006 + high_school + rgenrevpc + logNumpop + eopop + men_ineqincwag + grelimp + fedrepcons + urbanization_2 + logpop + inspop, data = AltBarber, index = c("st","year"), effect="twoways")
summary(testmodel02)

testmodel03 <- plm(CCI ~ jdiv + bindingonetermlimit + bindingtwotermlimit + rpcinc_annual_new + eousa + unemp + citi6006 + high_school + rgenrevpc + logNumpop + eopop + men_ineqincwag + grelimp + fedrepcons + urbanization_2 + logpop + inspop + ind_limits_2010*ind_limits_dummy + ind2pty_limits_2010*ind2pty_limits_dummy + pac_limits_2010*pac_limits_dummy, data = AltBarber, index = c("st","year"), effect="twoways")
summary(testmodel03)

testmodel05 <- plm(CCI ~ jdiv + bindingonetermlimit + bindingtwotermlimit + rpcinc_annual_new + eousa + unemp + citi6006 + high_school + rgenrevpc + logNumpop + eopop + men_ineqincwag + grelimp + urbanization_2 + logpop + inspop, data = AltBarber, index = c("st","year"), effect="twoways")
summary(testmodel05)

testmodel06 <- plm(CCI ~ jdiv + bindingonetermlimit + bindingtwotermlimit + rpcinc_annual_new + eousa + unemp + citi6006 + high_school + rgenrevpc + logNumpop + eopop + men_ineqincwag + grelimp + urbanization_2 + logpop + inspop + ind_limits_2010*ind_limits_dummy + ind2pty_limits_2010*ind2pty_limits_dummy + pac_limits_2010*pac_limits_dummy, data = AltBarber, index = c("st","year"), effect="twoways")
summary(testmodel06)

```