### 2018-05-23 MiP4.tex

\documentclass[]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdfauthor={Mark C. Hand},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{248,248,248}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\ImportTok}[1]{#1}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
\newcommand{\BuiltInTok}[1]{#1}
\newcommand{\ExtensionTok}[1]{#1}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
\newcommand{\RegionMarkerTok}[1]{#1}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{#1}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

%%% Use protect on footnotes to avoid problems with footnotes in titles
\let\rmarkdownfootnote\footnote%
\def\footnote{\protect\rmarkdownfootnote}

%%% Change title format to be more compact
\usepackage{titling}

% Create subtitle command for use in maketitle
\newcommand{\subtitle}[1]{
  \posttitle{
    \begin{center}\large#1\end{center}
    }
}

\setlength{\droptitle}{-2em}
  \title{Do campaign contribution limits limit corruption?\\
Evidence from the 50 states}
  \pretitle{\vspace{\droptitle}\centering\huge}
  \posttitle{\par}
  \author{Mark C. Hand}
  \preauthor{\centering\large\emph}
  \postauthor{\par}
  \predate{\centering\large\emph}
  \postdate{\par}
  \date{5/21/2018}

\begin{document}
\maketitle

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{knitr}\OperatorTok{::}\NormalTok{opts_chunk}\OperatorTok{$}\KeywordTok{set}\NormalTok{(}\DataTypeTok{echo =} \OtherTok{TRUE}\NormalTok{, }\DataTypeTok{cache =} \OtherTok{TRUE}\NormalTok{, }\DataTypeTok{message=}\OtherTok{FALSE}\NormalTok{, }\DataTypeTok{warning=}\OtherTok{FALSE}\NormalTok{)}
\KeywordTok{library}\NormalTok{(haven) }\CommentTok{# import .dta}
\KeywordTok{library}\NormalTok{(tidyverse) }\CommentTok{# everything }
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## -- Attaching packages ---------------------------------------------------- tidyverse 1.2.1 --
\end{verbatim}

\begin{verbatim}
## √ ggplot2 2.2.1     √ purrr   0.2.4
## √ tibble  1.4.2     √ dplyr   0.7.5
## √ tidyr   0.8.1     √ stringr 1.3.0
## √ readr   1.1.1     √ forcats 0.3.0
\end{verbatim}

\begin{verbatim}
## -- Conflicts ------------------------------------------------------- tidyverse_conflicts() --
## x dplyr::filter() masks stats::filter()
## x dplyr::lag()    masks stats::lag()
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{library}\NormalTok{(plm) }\CommentTok{# For linear panel regression}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## Loading required package: Formula
\end{verbatim}

\begin{verbatim}
## 
## Attaching package: 'plm'
\end{verbatim}

\begin{verbatim}
## The following objects are masked from 'package:dplyr':
## 
##     between, lag, lead
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# Alt and Lassen's "Enforcement and Public Corruption: Evidence from US States"}
\CommentTok{# http://epstein.wustl.edu/research/LPE.Alt.pdf}
\NormalTok{AltLassen <-}\StringTok{ }\KeywordTok{read_dta}\NormalTok{(}\StringTok{"data/Replication/AltLassenJLEO.dta"}\NormalTok{)}

\CommentTok{# Michael Barber's "Ideological Donors, Contribution Limits, and the Polarization of American Legislatures"  }
\CommentTok{# https://doi.org/10.1086/683453}
\NormalTok{Barber <-}\StringTok{ }\KeywordTok{read_csv}\NormalTok{(}\StringTok{"~/Documents/GitHub/Money-in-Politics/data/Barber/Barber_Limits_Data.csv"}\NormalTok{)}

\CommentTok{# Data from ICC }
\CommentTok{# www.greasethewheels.org/cci/}
\NormalTok{convictions <-}\StringTok{ }\KeywordTok{read_csv}\NormalTok{(}\StringTok{"~/Documents/GitHub/Money-in-Politics/data/ICS/CCI19762015.csv"}\NormalTok{)}
\NormalTok{reflections <-}\StringTok{ }\KeywordTok{read_csv}\NormalTok{(}\StringTok{"~/Documents/GitHub/Money-in-Politics/data/ICS/CRI19772013.csv"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# Combining Barber and ICS }
\NormalTok{corruption <-}\StringTok{ }\KeywordTok{full_join}\NormalTok{(convictions, reflections)}
\NormalTok{corruption <-}\StringTok{ }\NormalTok{corruption }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{rename}\NormalTok{(}\StringTok{"st"}\NormalTok{=}\StringTok{"State"}\NormalTok{, }\StringTok{"year"}\NormalTok{=}\StringTok{"Year"}\NormalTok{)}
\NormalTok{corruption}\OperatorTok{$}\NormalTok{st <-}\StringTok{ }\NormalTok{state.abb[}\KeywordTok{match}\NormalTok{(corruption}\OperatorTok{$}\NormalTok{st,state.name)]}
\NormalTok{BarberICC <-}\StringTok{ }\KeywordTok{left_join}\NormalTok{(Barber, corruption)}

\NormalTok{BarberICC <-}\StringTok{ }\NormalTok{BarberICC }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{mutate}\NormalTok{(}
    \DataTypeTok{ind_limits_dummy =} \KeywordTok{ifelse}\NormalTok{(ind_limits_}\DecValTok{2010}\OperatorTok{==}\DecValTok{99999}\NormalTok{,}\DecValTok{0}\NormalTok{,}\DecValTok{1}\NormalTok{),}
    \DataTypeTok{ind2pty_limits_dummy =} \KeywordTok{ifelse}\NormalTok{(ind2pty_limits_}\DecValTok{2010}\OperatorTok{==}\DecValTok{99999}\NormalTok{,}\DecValTok{0}\NormalTok{,}\DecValTok{1}\NormalTok{),}
    \DataTypeTok{pac_limits_dummy =} \KeywordTok{ifelse}\NormalTok{(pac_limits_}\DecValTok{2010}\OperatorTok{==}\DecValTok{99999}\NormalTok{,}\DecValTok{0}\NormalTok{,}\DecValTok{1}\NormalTok{),}
    \DataTypeTok{total_limits_dummy =} \KeywordTok{ifelse}\NormalTok{(total_limits_}\DecValTok{2010}\OperatorTok{==}\DecValTok{99999}\NormalTok{,}\DecValTok{0}\NormalTok{,}\DecValTok{1}\NormalTok{)}
\NormalTok{  )}

\CommentTok{# Adding AltLassen}
\CommentTok{# table(AltLassen$state_name)}
\NormalTok{AltLassen}\OperatorTok{$}\NormalTok{state_name[AltLassen}\OperatorTok{$}\NormalTok{state_name}\OperatorTok{==}\StringTok{"Californ"}\NormalTok{] =}\StringTok{ "California"}
\NormalTok{AltLassen}\OperatorTok{$}\NormalTok{state_name[AltLassen}\OperatorTok{$}\NormalTok{state_name}\OperatorTok{==}\StringTok{"Connecti"}\NormalTok{] =}\StringTok{ "Connecticut"}
\NormalTok{AltLassen}\OperatorTok{$}\NormalTok{state_name[AltLassen}\OperatorTok{$}\NormalTok{state_name}\OperatorTok{==}\StringTok{"Louisian"}\NormalTok{] =}\StringTok{ "Louisiana"}
\NormalTok{AltLassen}\OperatorTok{$}\NormalTok{state_name[AltLassen}\OperatorTok{$}\NormalTok{state_name}\OperatorTok{==}\StringTok{"Massachu"}\NormalTok{] =}\StringTok{ "Massachusetts"}
\NormalTok{AltLassen}\OperatorTok{$}\NormalTok{state_name[AltLassen}\OperatorTok{$}\NormalTok{state_name}\OperatorTok{==}\StringTok{"Minnesot"}\NormalTok{] =}\StringTok{ "Minnesota"}
\NormalTok{AltLassen}\OperatorTok{$}\NormalTok{state_name[AltLassen}\OperatorTok{$}\NormalTok{state_name}\OperatorTok{==}\StringTok{"Mississi"}\NormalTok{] =}\StringTok{ "Mississippi"}
\NormalTok{AltLassen}\OperatorTok{$}\NormalTok{state_name[AltLassen}\OperatorTok{$}\NormalTok{state_name}\OperatorTok{==}\StringTok{"NewHamps"}\NormalTok{] =}\StringTok{ "New Hampshire"}
\NormalTok{AltLassen}\OperatorTok{$}\NormalTok{state_name[AltLassen}\OperatorTok{$}\NormalTok{state_name}\OperatorTok{==}\StringTok{"NewJerse"}\NormalTok{] =}\StringTok{ "New Jersey"}
\NormalTok{AltLassen}\OperatorTok{$}\NormalTok{state_name[AltLassen}\OperatorTok{$}\NormalTok{state_name}\OperatorTok{==}\StringTok{"NewMexic"}\NormalTok{] =}\StringTok{ "New Mexico"}
\NormalTok{AltLassen}\OperatorTok{$}\NormalTok{state_name[AltLassen}\OperatorTok{$}\NormalTok{state_name}\OperatorTok{==}\StringTok{"NewYork"}\NormalTok{] =}\StringTok{ "New York"}
\NormalTok{AltLassen}\OperatorTok{$}\NormalTok{state_name[AltLassen}\OperatorTok{$}\NormalTok{state_name}\OperatorTok{==}\StringTok{"NorthCar"}\NormalTok{] =}\StringTok{ "North Carolina"}
\NormalTok{AltLassen}\OperatorTok{$}\NormalTok{state_name[AltLassen}\OperatorTok{$}\NormalTok{state_name}\OperatorTok{==}\StringTok{"NorthDak"}\NormalTok{] =}\StringTok{ "North Dakota"}
\NormalTok{AltLassen}\OperatorTok{$}\NormalTok{state_name[AltLassen}\OperatorTok{$}\NormalTok{state_name}\OperatorTok{==}\StringTok{"Pennsylv"}\NormalTok{] =}\StringTok{ "Pennsylvania"}
\NormalTok{AltLassen}\OperatorTok{$}\NormalTok{state_name[AltLassen}\OperatorTok{$}\NormalTok{state_name}\OperatorTok{==}\StringTok{"RhodeIsl"}\NormalTok{] =}\StringTok{ "Rhode Island"}
\NormalTok{AltLassen}\OperatorTok{$}\NormalTok{state_name[AltLassen}\OperatorTok{$}\NormalTok{state_name}\OperatorTok{==}\StringTok{"SouthCar"}\NormalTok{] =}\StringTok{ "South Carolina"}
\NormalTok{AltLassen}\OperatorTok{$}\NormalTok{state_name[AltLassen}\OperatorTok{$}\NormalTok{state_name}\OperatorTok{==}\StringTok{"SouthDak"}\NormalTok{] =}\StringTok{ "South Dakota"}
\NormalTok{AltLassen}\OperatorTok{$}\NormalTok{state_name[AltLassen}\OperatorTok{$}\NormalTok{state_name}\OperatorTok{==}\StringTok{"Tennesse"}\NormalTok{] =}\StringTok{ "Tennessee"}
\NormalTok{AltLassen}\OperatorTok{$}\NormalTok{state_name[AltLassen}\OperatorTok{$}\NormalTok{state_name}\OperatorTok{==}\StringTok{"Washingt"}\NormalTok{] =}\StringTok{ "Washington"}
\NormalTok{AltLassen}\OperatorTok{$}\NormalTok{state_name[AltLassen}\OperatorTok{$}\NormalTok{state_name}\OperatorTok{==}\StringTok{"WestVirg"}\NormalTok{] =}\StringTok{ "West Virginia"}
\NormalTok{AltLassen}\OperatorTok{$}\NormalTok{state_name[AltLassen}\OperatorTok{$}\NormalTok{state_name}\OperatorTok{==}\StringTok{"Wisconsi"}\NormalTok{] =}\StringTok{ "Wisconsin"}
\NormalTok{AltLassen}\OperatorTok{$}\NormalTok{state_name <-}\StringTok{ }\NormalTok{state.abb[}\KeywordTok{match}\NormalTok{(AltLassen}\OperatorTok{$}\NormalTok{state_name,state.name)]}
\NormalTok{AltLassen <-}\StringTok{ }\NormalTok{AltLassen }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{rename}\NormalTok{(}\StringTok{"st"}\NormalTok{=}\StringTok{"state_name"}\NormalTok{)}
\NormalTok{AltLassen}\OperatorTok{$}\NormalTok{year <-}\KeywordTok{as.integer}\NormalTok{(AltLassen}\OperatorTok{$}\NormalTok{year)}

\NormalTok{AltBarber <-}\StringTok{ }\KeywordTok{inner_join}\NormalTok{(BarberICC, AltLassen)}
\end{Highlighting}
\end{Shaded}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{#Next: Panel model of corruption against unemployment and high school? }
\NormalTok{testmodel01 <-}\StringTok{ }\KeywordTok{plm}\NormalTok{(CCI }\OperatorTok{~}\StringTok{ }\NormalTok{rpcinc_annual_new }\OperatorTok{+}\StringTok{ }\NormalTok{unemp }\OperatorTok{+}\StringTok{ }\NormalTok{high_school }\OperatorTok{+}\StringTok{ }\NormalTok{logNumpop }\OperatorTok{+}\StringTok{ }\NormalTok{citi6006 }\OperatorTok{+}\StringTok{ }\NormalTok{grelimp, }\DataTypeTok{data =}\NormalTok{ AltBarber, }\DataTypeTok{index =} \KeywordTok{c}\NormalTok{(}\StringTok{"st"}\NormalTok{,}\StringTok{"year"}\NormalTok{), }\DataTypeTok{effect=}\StringTok{"twoways"}\NormalTok{)}
\KeywordTok{summary}\NormalTok{(testmodel01)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## Twoways effects Within Model
## 
## Call:
## plm(formula = CCI ~ rpcinc_annual_new + unemp + high_school + 
##     logNumpop + citi6006 + grelimp, data = AltBarber, effect = "twoways", 
##     index = c("st", "year"))
## 
## Unbalanced Panel: n = 48, T = 13-15, N = 713
## 
## Residuals:
##     Min.  1st Qu.   Median  3rd Qu.     Max. 
## -4.81797 -1.00621 -0.18884  0.81311 13.77887 
## 
## Coefficients:
##                      Estimate  Std. Error t-value Pr(>|t|)    
## rpcinc_annual_new -4.0754e-04  1.8335e-04 -2.2228  0.02658 *  
## unemp             -6.3066e-02  1.2605e-01 -0.5003  0.61703    
## high_school       -1.2978e+00  5.5780e+00 -0.2327  0.81610    
## logNumpop          2.2137e-01  1.6150e-02 13.7072  < 2e-16 ***
## citi6006           2.1570e-03  1.6662e-02  0.1295  0.89704    
## grelimp           -1.1922e+01  3.6482e+01 -0.3268  0.74393    
## ---
## Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1
## 
## Total Sum of Squares:    3132
## Residual Sum of Squares: 2404
## R-Squared:      0.23244
## Adj. R-Squared: 0.1527
## F-statistic: 32.5533 on 6 and 645 DF, p-value: < 2.22e-16
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{testmodel02 <-}\StringTok{ }\KeywordTok{plm}\NormalTok{(CCI }\OperatorTok{~}\StringTok{ }\NormalTok{jdiv }\OperatorTok{+}\StringTok{ }\NormalTok{bindingonetermlimit }\OperatorTok{+}\StringTok{ }\NormalTok{bindingtwotermlimit }\OperatorTok{+}\StringTok{ }\NormalTok{rpcinc_annual_new }\OperatorTok{+}\StringTok{ }\NormalTok{eousa }\OperatorTok{+}\StringTok{ }\NormalTok{unemp }\OperatorTok{+}\StringTok{ }\NormalTok{citi6006 }\OperatorTok{+}\StringTok{ }\NormalTok{high_school }\OperatorTok{+}\StringTok{ }\NormalTok{rgenrevpc }\OperatorTok{+}\StringTok{ }\NormalTok{logNumpop }\OperatorTok{+}\StringTok{ }\NormalTok{eopop }\OperatorTok{+}\StringTok{ }\NormalTok{men_ineqincwag }\OperatorTok{+}\StringTok{ }\NormalTok{grelimp }\OperatorTok{+}\StringTok{ }\NormalTok{fedrepcons }\OperatorTok{+}\StringTok{ }\NormalTok{urbanization_}\DecValTok{2} \OperatorTok{+}\StringTok{ }\NormalTok{logpop }\OperatorTok{+}\StringTok{ }\NormalTok{inspop, }\DataTypeTok{data =}\NormalTok{ AltBarber, }\DataTypeTok{index =} \KeywordTok{c}\NormalTok{(}\StringTok{"st"}\NormalTok{,}\StringTok{"year"}\NormalTok{), }\DataTypeTok{effect=}\StringTok{"twoways"}\NormalTok{)}
\KeywordTok{summary}\NormalTok{(testmodel02)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## Twoways effects Within Model
## 
## Call:
## plm(formula = CCI ~ jdiv + bindingonetermlimit + bindingtwotermlimit + 
##     rpcinc_annual_new + eousa + unemp + citi6006 + high_school + 
##     rgenrevpc + logNumpop + eopop + men_ineqincwag + grelimp + 
##     fedrepcons + urbanization_2 + logpop + inspop, data = AltBarber, 
##     effect = "twoways", index = c("st", "year"))
## 
## Unbalanced Panel: n = 48, T = 11-13, N = 618
## 
## Residuals:
##     Min.  1st Qu.   Median  3rd Qu.     Max. 
## -5.25732 -1.06561 -0.12498  0.83593 13.44049 
## 
## Coefficients:
##                        Estimate  Std. Error t-value  Pr(>|t|)    
## jdiv                -0.20190256  0.20282511 -0.9955  0.319962    
## bindingonetermlimit -0.39118843  1.27029686 -0.3080  0.758239    
## bindingtwotermlimit -0.04300755  0.22054353 -0.1950  0.845461    
## rpcinc_annual_new   -0.00056598  0.00025004 -2.2636  0.023995 *  
## eousa                0.00535221  0.01069341  0.5005  0.616916    
## unemp               -0.10017217  0.15209202 -0.6586  0.510414    
## citi6006            -0.01032798  0.01913966 -0.5396  0.589687    
## high_school         -5.14098696  7.01773427 -0.7326  0.464138    
## rgenrevpc            0.00041420  0.00112210  0.3691  0.712178    
## logNumpop            0.22566553  0.01828559 12.3412 < 2.2e-16 ***
## eopop                0.01548702  0.06500440  0.2382  0.811781    
## men_ineqincwag       0.13088787  0.04447225  2.9431  0.003388 ** 
## grelimp             22.11759146 43.75544159  0.5055  0.613426    
## fedrepcons          -0.06259835  0.02282555 -2.7425  0.006300 ** 
## urbanization_2       0.14630029  0.10887078  1.3438  0.179577    
## logpop              -3.05635200  3.16861625 -0.9646  0.335191    
## inspop              -0.00162197  0.00226241 -0.7169  0.473732    
## ---
## Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1
## 
## Total Sum of Squares:    2911.7
## Residual Sum of Squares: 2136.5
## R-Squared:      0.26624
## Adj. R-Squared: 0.16316
## F-statistic: 11.5467 on 17 and 541 DF, p-value: < 2.22e-16
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{testmodel03 <-}\StringTok{ }\KeywordTok{plm}\NormalTok{(CCI }\OperatorTok{~}\StringTok{ }\NormalTok{jdiv }\OperatorTok{+}\StringTok{ }\NormalTok{bindingonetermlimit }\OperatorTok{+}\StringTok{ }\NormalTok{bindingtwotermlimit }\OperatorTok{+}\StringTok{ }\NormalTok{rpcinc_annual_new }\OperatorTok{+}\StringTok{ }\NormalTok{eousa }\OperatorTok{+}\StringTok{ }\NormalTok{unemp }\OperatorTok{+}\StringTok{ }\NormalTok{citi6006 }\OperatorTok{+}\StringTok{ }\NormalTok{high_school }\OperatorTok{+}\StringTok{ }\NormalTok{rgenrevpc }\OperatorTok{+}\StringTok{ }\NormalTok{logNumpop }\OperatorTok{+}\StringTok{ }\NormalTok{eopop }\OperatorTok{+}\StringTok{ }\NormalTok{men_ineqincwag }\OperatorTok{+}\StringTok{ }\NormalTok{grelimp }\OperatorTok{+}\StringTok{ }\NormalTok{fedrepcons }\OperatorTok{+}\StringTok{ }\NormalTok{urbanization_}\DecValTok{2} \OperatorTok{+}\StringTok{ }\NormalTok{logpop }\OperatorTok{+}\StringTok{ }\NormalTok{inspop }\OperatorTok{+}\StringTok{ }\NormalTok{ind_limits_}\DecValTok{2010}\OperatorTok{*}\NormalTok{ind_limits_dummy }\OperatorTok{+}\StringTok{ }\NormalTok{ind2pty_limits_}\DecValTok{2010}\OperatorTok{*}\NormalTok{ind2pty_limits_dummy }\OperatorTok{+}\StringTok{ }\NormalTok{pac_limits_}\DecValTok{2010}\OperatorTok{*}\NormalTok{pac_limits_dummy, }\DataTypeTok{data =}\NormalTok{ AltBarber, }\DataTypeTok{index =} \KeywordTok{c}\NormalTok{(}\StringTok{"st"}\NormalTok{,}\StringTok{"year"}\NormalTok{), }\DataTypeTok{effect=}\StringTok{"twoways"}\NormalTok{)}
\KeywordTok{summary}\NormalTok{(testmodel03)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## Twoways effects Within Model
## 
## Call:
## plm(formula = CCI ~ jdiv + bindingonetermlimit + bindingtwotermlimit + 
##     rpcinc_annual_new + eousa + unemp + citi6006 + high_school + 
##     rgenrevpc + logNumpop + eopop + men_ineqincwag + grelimp + 
##     fedrepcons + urbanization_2 + logpop + inspop + ind_limits_2010 * 
##     ind_limits_dummy + ind2pty_limits_2010 * ind2pty_limits_dummy + 
##     pac_limits_2010 * pac_limits_dummy, data = AltBarber, effect = "twoways", 
##     index = c("st", "year"))
## 
## Unbalanced Panel: n = 48, T = 11-13, N = 618
## 
## Residuals:
##     Min.  1st Qu.   Median  3rd Qu.     Max. 
## -5.22029 -1.08852 -0.12102  0.83586 13.38407 
## 
## Coefficients: (3 dropped because of singularities)
##                         Estimate  Std. Error t-value  Pr(>|t|)    
## jdiv                 -1.7767e-01  2.0591e-01 -0.8628  0.388617    
## bindingonetermlimit  -7.2323e-01  1.3644e+00 -0.5301  0.596282    
## bindingtwotermlimit  -1.8335e-02  2.2332e-01 -0.0821  0.934595    
## rpcinc_annual_new    -5.8467e-04  2.5598e-04 -2.2840  0.022761 *  
## eousa                 5.5245e-03  1.0849e-02  0.5092  0.610809    
## unemp                -1.0827e-01  1.5814e-01 -0.6846  0.493881    
## citi6006             -7.6448e-03  1.9416e-02 -0.3937  0.693931    
## high_school          -4.5817e+00  7.2155e+00 -0.6350  0.525707    
## rgenrevpc             6.8401e-04  1.1443e-03  0.5978  0.550251    
## logNumpop             2.2471e-01  1.8497e-02 12.1486 < 2.2e-16 ***
## eopop                 2.1830e-02  6.6799e-02  0.3268  0.743944    
## men_ineqincwag        1.3021e-01  4.5213e-02  2.8800  0.004137 ** 
## grelimp               2.4629e+01  4.6362e+01  0.5312  0.595488    
## fedrepcons           -6.4202e-02  2.3003e-02 -2.7910  0.005442 ** 
## urbanization_2        1.4034e-01  1.1000e-01  1.2758  0.202593    
## logpop               -3.8183e+00  3.4541e+00 -1.1054  0.269467    
## inspop               -1.2946e-03  2.4236e-03 -0.5342  0.593442    
## ind_limits_2010       4.6294e-05  8.2230e-05  0.5630  0.573685    
## ind_limits_dummy      4.2620e+00  7.9863e+00  0.5337  0.593794    
## ind2pty_limits_2010   3.5794e-06  6.6072e-06  0.5417  0.588223    
## ind2pty_limits_dummy  1.8875e-02  7.0763e-01  0.0267  0.978730    
## pac_limits_2010      -6.8521e-06  3.3117e-05 -0.2069  0.836160    
## pac_limits_dummy     -8.1158e-02  2.9465e+00 -0.0275  0.978036    
## ---
## Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1
## 
## Total Sum of Squares:    2911.7
## Residual Sum of Squares: 2128.6
## R-Squared:      0.26896
## Adj. R-Squared: 0.15692
## F-statistic: 8.55816 on 23 and 535 DF, p-value: < 2.22e-16
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{testmodel05 <-}\StringTok{ }\KeywordTok{plm}\NormalTok{(CCI }\OperatorTok{~}\StringTok{ }\NormalTok{jdiv }\OperatorTok{+}\StringTok{ }\NormalTok{bindingonetermlimit }\OperatorTok{+}\StringTok{ }\NormalTok{bindingtwotermlimit }\OperatorTok{+}\StringTok{ }\NormalTok{rpcinc_annual_new }\OperatorTok{+}\StringTok{ }\NormalTok{eousa }\OperatorTok{+}\StringTok{ }\NormalTok{unemp }\OperatorTok{+}\StringTok{ }\NormalTok{citi6006 }\OperatorTok{+}\StringTok{ }\NormalTok{high_school }\OperatorTok{+}\StringTok{ }\NormalTok{rgenrevpc }\OperatorTok{+}\StringTok{ }\NormalTok{logNumpop }\OperatorTok{+}\StringTok{ }\NormalTok{eopop }\OperatorTok{+}\StringTok{ }\NormalTok{men_ineqincwag }\OperatorTok{+}\StringTok{ }\NormalTok{grelimp }\OperatorTok{+}\StringTok{ }\NormalTok{urbanization_}\DecValTok{2} \OperatorTok{+}\StringTok{ }\NormalTok{logpop }\OperatorTok{+}\StringTok{ }\NormalTok{inspop, }\DataTypeTok{data =}\NormalTok{ AltBarber, }\DataTypeTok{index =} \KeywordTok{c}\NormalTok{(}\StringTok{"st"}\NormalTok{,}\StringTok{"year"}\NormalTok{), }\DataTypeTok{effect=}\StringTok{"twoways"}\NormalTok{)}
\KeywordTok{summary}\NormalTok{(testmodel05)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## Twoways effects Within Model
## 
## Call:
## plm(formula = CCI ~ jdiv + bindingonetermlimit + bindingtwotermlimit + 
##     rpcinc_annual_new + eousa + unemp + citi6006 + high_school + 
##     rgenrevpc + logNumpop + eopop + men_ineqincwag + grelimp + 
##     urbanization_2 + logpop + inspop, data = AltBarber, effect = "twoways", 
##     index = c("st", "year"))
## 
## Unbalanced Panel: n = 48, T = 11-13, N = 618
## 
## Residuals:
##     Min.  1st Qu.   Median  3rd Qu.     Max. 
## -5.19413 -1.07814 -0.11191  0.84387 13.69793 
## 
## Coefficients:
##                        Estimate  Std. Error t-value  Pr(>|t|)    
## jdiv                -0.17829000  0.20385769 -0.8746  0.382189    
## bindingonetermlimit -0.37832749  1.27790716 -0.2961  0.767303    
## bindingtwotermlimit -0.10234395  0.22079607 -0.4635  0.643176    
## rpcinc_annual_new   -0.00059742  0.00025127 -2.3776  0.017772 *  
## eousa                0.00496122  0.01075659  0.4612  0.644821    
## unemp               -0.12961622  0.15262255 -0.8493  0.396112    
## citi6006            -0.00882295  0.01924654 -0.4584  0.646837    
## high_school         -4.43957589  7.05513532 -0.6293  0.529438    
## rgenrevpc            0.00057200  0.00112735  0.5074  0.612089    
## logNumpop            0.22637491  0.01839343 12.3074 < 2.2e-16 ***
## eopop                0.02908226  0.06520384  0.4460  0.655760    
## men_ineqincwag       0.12693924  0.04471553  2.8388  0.004698 ** 
## grelimp             13.18285583 43.89569361  0.3003  0.764046    
## urbanization_2       0.15342180  0.10949261  1.4012  0.161725    
## logpop              -2.40376676  3.17861979 -0.7562  0.449840    
## inspop              -0.00145727  0.00227518 -0.6405  0.522112    
## ---
## Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1
## 
## Total Sum of Squares:    2911.7
## Residual Sum of Squares: 2166.2
## R-Squared:      0.25603
## Adj. R-Squared: 0.15309
## F-statistic: 11.6581 on 16 and 542 DF, p-value: < 2.22e-16
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{testmodel06 <-}\StringTok{ }\KeywordTok{plm}\NormalTok{(CCI }\OperatorTok{~}\StringTok{ }\NormalTok{jdiv }\OperatorTok{+}\StringTok{ }\NormalTok{bindingonetermlimit }\OperatorTok{+}\StringTok{ }\NormalTok{bindingtwotermlimit }\OperatorTok{+}\StringTok{ }\NormalTok{rpcinc_annual_new }\OperatorTok{+}\StringTok{ }\NormalTok{eousa }\OperatorTok{+}\StringTok{ }\NormalTok{unemp }\OperatorTok{+}\StringTok{ }\NormalTok{citi6006 }\OperatorTok{+}\StringTok{ }\NormalTok{high_school }\OperatorTok{+}\StringTok{ }\NormalTok{rgenrevpc }\OperatorTok{+}\StringTok{ }\NormalTok{logNumpop }\OperatorTok{+}\StringTok{ }\NormalTok{eopop }\OperatorTok{+}\StringTok{ }\NormalTok{men_ineqincwag }\OperatorTok{+}\StringTok{ }\NormalTok{grelimp }\OperatorTok{+}\StringTok{ }\NormalTok{urbanization_}\DecValTok{2} \OperatorTok{+}\StringTok{ }\NormalTok{logpop }\OperatorTok{+}\StringTok{ }\NormalTok{inspop }\OperatorTok{+}\StringTok{ }\NormalTok{ind_limits_}\DecValTok{2010}\OperatorTok{*}\NormalTok{ind_limits_dummy }\OperatorTok{+}\StringTok{ }\NormalTok{ind2pty_limits_}\DecValTok{2010}\OperatorTok{*}\NormalTok{ind2pty_limits_dummy }\OperatorTok{+}\StringTok{ }\NormalTok{pac_limits_}\DecValTok{2010}\OperatorTok{*}\NormalTok{pac_limits_dummy, }\DataTypeTok{data =}\NormalTok{ AltBarber, }\DataTypeTok{index =} \KeywordTok{c}\NormalTok{(}\StringTok{"st"}\NormalTok{,}\StringTok{"year"}\NormalTok{), }\DataTypeTok{effect=}\StringTok{"twoways"}\NormalTok{)}
\KeywordTok{summary}\NormalTok{(testmodel06)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## Twoways effects Within Model
## 
## Call:
## plm(formula = CCI ~ jdiv + bindingonetermlimit + bindingtwotermlimit + 
##     rpcinc_annual_new + eousa + unemp + citi6006 + high_school + 
##     rgenrevpc + logNumpop + eopop + men_ineqincwag + grelimp + 
##     urbanization_2 + logpop + inspop + ind_limits_2010 * ind_limits_dummy + 
##     ind2pty_limits_2010 * ind2pty_limits_dummy + pac_limits_2010 * 
##     pac_limits_dummy, data = AltBarber, effect = "twoways", index = c("st", 
##     "year"))
## 
## Unbalanced Panel: n = 48, T = 11-13, N = 618
## 
## Residuals:
##     Min.  1st Qu.   Median  3rd Qu.     Max. 
## -5.15526 -1.04988 -0.11867  0.84917 13.64210 
## 
## Coefficients: (3 dropped because of singularities)
##                         Estimate  Std. Error t-value Pr(>|t|)    
## jdiv                 -1.5601e-01  2.0706e-01 -0.7534   0.4515    
## bindingonetermlimit  -6.9102e-01  1.3730e+00 -0.5033   0.6150    
## bindingtwotermlimit  -8.3880e-02  2.2348e-01 -0.3753   0.7076    
## rpcinc_annual_new    -6.1889e-04  2.5730e-04 -2.4053   0.0165 *  
## eousa                 4.9832e-03  1.0916e-02  0.4565   0.6482    
## unemp                -1.4377e-01  1.5862e-01 -0.9064   0.3652    
## citi6006             -6.0887e-03  1.9530e-02 -0.3118   0.7553    
## high_school          -3.7101e+00  7.2542e+00 -0.5114   0.6093    
## rgenrevpc             8.2850e-04  1.1503e-03  0.7202   0.4717    
## logNumpop             2.2595e-01  1.8608e-02 12.1423   <2e-16 ***
## eopop                 3.3961e-02  6.7078e-02  0.5063   0.6129    
## men_ineqincwag        1.2678e-01  4.5481e-02  2.7875   0.0055 ** 
## grelimp               1.6483e+01  4.6563e+01  0.3540   0.7235    
## urbanization_2        1.4689e-01  1.1067e-01  1.3272   0.1850    
## logpop               -3.1599e+00  3.4678e+00 -0.9112   0.3626    
## inspop               -1.0530e-03  2.4374e-03 -0.4320   0.6659    
## ind_limits_2010       4.1554e-05  8.2732e-05  0.5023   0.6157    
## ind_limits_dummy      3.8560e+00  8.0354e+00  0.4799   0.6315    
## ind2pty_limits_2010   3.5071e-06  6.6489e-06  0.5275   0.5981    
## ind2pty_limits_dummy -4.3971e-02  7.1174e-01 -0.0618   0.9508    
## pac_limits_2010      -2.2065e-06  3.3283e-05 -0.0663   0.9472    
## pac_limits_dummy      2.5907e-01  2.9626e+00  0.0874   0.9303    
## ---
## Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1
## 
## Total Sum of Squares:    2911.7
## Residual Sum of Squares: 2159.6
## R-Squared:      0.25832
## Adj. R-Squared: 0.14624
## F-statistic: 8.48559 on 22 and 536 DF, p-value: < 2.22e-16
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{BarberICC }\OperatorTok{%>%}\StringTok{ }\KeywordTok{arrange}\NormalTok{(year) }\OperatorTok{%>%}\StringTok{ }\KeywordTok{group_by}\NormalTok{(st) }\OperatorTok{%>%}\StringTok{ }\KeywordTok{mutate}\NormalTok{(}
  \DataTypeTok{ind_d =}\NormalTok{ (ind_limits_}\DecValTok{2010} \OperatorTok{-}\StringTok{ }\NormalTok{dplyr}\OperatorTok{::}\KeywordTok{lag}\NormalTok{(ind_limits_}\DecValTok{2010}\NormalTok{)),}
  \DataTypeTok{ind_dummy_d =}\NormalTok{ (ind_limits_dummy }\OperatorTok{-}\StringTok{ }\NormalTok{dplyr}\OperatorTok{::}\KeywordTok{lag}\NormalTok{(ind_limits_dummy)),}
  \DataTypeTok{pac_d =}\NormalTok{ (pac_limits_}\DecValTok{2010} \OperatorTok{-}\StringTok{ }\NormalTok{dplyr}\OperatorTok{::}\KeywordTok{lag}\NormalTok{(pac_limits_}\DecValTok{2010}\NormalTok{)),}
  \DataTypeTok{pac_dummy_d =}\NormalTok{ (pac_limits_dummy }\OperatorTok{-}\StringTok{ }\NormalTok{dplyr}\OperatorTok{::}\KeywordTok{lag}\NormalTok{(pac_limits_dummy)),}
  \DataTypeTok{ind2pty_d =}\NormalTok{ (ind2pty_limits_}\DecValTok{2010} \OperatorTok{-}\StringTok{ }\NormalTok{dplyr}\OperatorTok{::}\KeywordTok{lag}\NormalTok{(ind2pty_limits_}\DecValTok{2010}\NormalTok{)),}
  \DataTypeTok{ind2pty_dummy_d =}\NormalTok{ (ind2pty_limits_dummy }\OperatorTok{-}\StringTok{ }\NormalTok{dplyr}\OperatorTok{::}\KeywordTok{lag}\NormalTok{(ind2pty_limits_dummy))}
\NormalTok{  ) ->}\StringTok{ }\NormalTok{testframe}

\NormalTok{testframe }\OperatorTok{%>%}\StringTok{ }\KeywordTok{select}\NormalTok{(}\DecValTok{1}\OperatorTok{:}\DecValTok{2}\NormalTok{,}\DecValTok{8}\NormalTok{,}\DecValTok{11}\NormalTok{,}\DecValTok{3}\OperatorTok{:}\DecValTok{7}\NormalTok{,}\DecValTok{12}\OperatorTok{:}\DecValTok{21}\NormalTok{) ->}\StringTok{ }\NormalTok{DiDsmall }
\KeywordTok{summary}\NormalTok{(DiDsmall)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##       st                 year           CCI              CRI        
##  Length:1173        Min.   :1990   Min.   : 0.000   Min.   :0.0000  
##  Class :character   1st Qu.:1995   1st Qu.: 1.348   1st Qu.:0.1577  
##  Mode  :character   Median :2001   Median : 2.562   Median :0.2585  
##                     Mean   :2001   Mean   : 3.226   Mean   :0.2911  
##                     3rd Qu.:2007   3rd Qu.: 4.296   3rd Qu.:0.3862  
##                     Max.   :2012   Max.   :33.846   Max.   :2.1587  
##                                    NA's   :24       NA's   :23      
##  ind_limits_2010 ind2pty_limits_2010 pty_limits_2010  pac_limits_2010
##  Min.   :  143   Min.   :   834      Min.   :   246   Min.   :  223  
##  1st Qu.: 1397   1st Qu.: 24238      1st Qu.: 11707   1st Qu.: 2103  
##  Median : 4310   Median : 99999      Median : 99999   Median :10128  
##  Mean   :32911   Mean   : 72592      Mean   : 64022   Mean   :39604  
##  3rd Qu.:99999   3rd Qu.: 99999      3rd Qu.: 99999   3rd Qu.:99999  
##  Max.   :99999   Max.   :196470      Max.   :150904   Max.   :99999  
##  NA's   :23      NA's   :23          NA's   :23       NA's   :23     
##  total_limits_2010 ind_limits_dummy ind2pty_limits_dummy pac_limits_dummy
##  Min.   :  4772    Min.   :0.0000   Min.   :0.0000       Min.   :0.0000  
##  1st Qu.: 99999    1st Qu.:0.0000   1st Qu.:0.0000       1st Qu.:0.0000  
##  Median : 99999    Median :1.0000   Median :0.0000       Median :1.0000  
##  Mean   : 96196    Mean   :0.6974   Mean   :0.4104       Mean   :0.6452  
##  3rd Qu.: 99999    3rd Qu.:1.0000   3rd Qu.:1.0000       3rd Qu.:1.0000  
##  Max.   :500511    Max.   :1.0000   Max.   :1.0000       Max.   :1.0000  
##  NA's   :23        NA's   :23       NA's   :23           NA's   :23      
##  total_limits_dummy     ind_d           ind_dummy_d      
##  Min.   :0.0000     Min.   :-99856.0   Min.   :-1.00000  
##  1st Qu.:0.0000     1st Qu.:   -77.0   1st Qu.: 0.00000  
##  Median :0.0000     Median :   -20.0   Median : 0.00000  
##  Mean   :0.1635     Mean   :  -913.4   Mean   : 0.00909  
##  3rd Qu.:0.0000     3rd Qu.:     0.0   3rd Qu.: 0.00000  
##  Max.   :1.0000     Max.   : 99731.0   Max.   : 1.00000  
##  NA's   :23         NA's   :73         NA's   :73        
##      pac_d           pac_dummy_d         ind2pty_d        
##  Min.   :-99713.0   Min.   :-1.00000   Min.   :-115296.0  
##  1st Qu.:  -113.0   1st Qu.: 0.00000   1st Qu.:   -263.2  
##  Median :   -19.5   Median : 0.00000   Median :      0.0  
##  Mean   : -1194.4   Mean   : 0.01364   Mean   :   -533.7  
##  3rd Qu.:     0.0   3rd Qu.: 0.00000   3rd Qu.:      0.0  
##  Max.   : 99721.0   Max.   : 1.00000   Max.   : 120086.0  
##  NA's   :73         NA's   :73         NA's   :73         
##  ind2pty_dummy_d   
##  Min.   :-1.00000  
##  1st Qu.: 0.00000  
##  Median : 0.00000  
##  Mean   : 0.00818  
##  3rd Qu.: 0.00000  
##  Max.   : 1.00000  
##  NA's   :73
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# Now I have what I think are three multi-level categorical independent variables. }
\CommentTok{# https://stats.idre.ucla.edu/r/modules/coding-for-categorical-variables-in-regression-models/}
\CommentTok{#1. When limits were imposed (1) or went down, did corruption go down? }
\CommentTok{#2. When limits were removed (-1) or went up, did corruption go up? }

\NormalTok{DiDsmall}\OperatorTok{$}\NormalTok{ind_dummy_f <-}\StringTok{ }\KeywordTok{factor}\NormalTok{(DiDsmall}\OperatorTok{$}\NormalTok{ind_dummy_d)}
\NormalTok{DiDsmall <-}\StringTok{ }\KeywordTok{within}\NormalTok{(DiDsmall, ind_dummy_f <-}\StringTok{ }\KeywordTok{relevel}\NormalTok{(ind_dummy_f, }\DataTypeTok{ref =} \StringTok{"0"}\NormalTok{))}
\NormalTok{DiDsmall}\OperatorTok{$}\NormalTok{pac_dummy_f <-}\StringTok{ }\KeywordTok{factor}\NormalTok{(DiDsmall}\OperatorTok{$}\NormalTok{pac_dummy_d)}
\NormalTok{DiDsmall <-}\StringTok{ }\KeywordTok{within}\NormalTok{(DiDsmall, pac_dummy_f <-}\StringTok{ }\KeywordTok{relevel}\NormalTok{(pac_dummy_f, }\DataTypeTok{ref =} \StringTok{"0"}\NormalTok{))}
\NormalTok{DiDsmall}\OperatorTok{$}\NormalTok{ind2pty_dummy_f <-}\StringTok{ }\KeywordTok{factor}\NormalTok{(DiDsmall}\OperatorTok{$}\NormalTok{ind2pty_dummy_d)}
\NormalTok{DiDsmall <-}\StringTok{ }\KeywordTok{within}\NormalTok{(DiDsmall, ind2pty_dummy_f <-}\StringTok{ }\KeywordTok{relevel}\NormalTok{(ind2pty_dummy_f, }\DataTypeTok{ref =} \StringTok{"0"}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{summary}\NormalTok{(}\KeywordTok{lm}\NormalTok{(CCI }\OperatorTok{~}\StringTok{ }\NormalTok{ind_dummy_f, }\DataTypeTok{data =}\NormalTok{ DiDsmall))}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## 
## Call:
## lm(formula = CCI ~ ind_dummy_f, data = DiDsmall)
## 
## Residuals:
##     Min      1Q  Median      3Q     Max 
## -3.1979 -1.8494 -0.6355  1.0862 30.6483 
## 
## Coefficients:
##               Estimate Std. Error t value Pr(>|t|)    
## (Intercept)    3.19792    0.08943  35.758   <2e-16 ***
## ind_dummy_f-1 -0.73316    1.20153  -0.610    0.542    
## ind_dummy_f1  -0.46753    0.73917  -0.633    0.527    
## ---
## Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1
## 
## Residual standard error: 2.935 on 1096 degrees of freedom
##   (74 observations deleted due to missingness)
## Multiple R-squared:  0.000698,   Adjusted R-squared:  -0.001126 
## F-statistic: 0.3827 on 2 and 1096 DF,  p-value: 0.6821
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{summary}\NormalTok{(}\KeywordTok{lm}\NormalTok{(CCI }\OperatorTok{~}\StringTok{ }\NormalTok{pac_dummy_f, }\DataTypeTok{data =}\NormalTok{ DiDsmall))}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## 
## Call:
## lm(formula = CCI ~ pac_dummy_f, data = DiDsmall)
## 
## Residuals:
##     Min      1Q  Median      3Q     Max 
## -3.1969 -1.8484 -0.6345  1.0814 30.6493 
## 
## Coefficients:
##               Estimate Std. Error t value Pr(>|t|)    
## (Intercept)    3.19693    0.08965  35.659   <2e-16 ***
## pac_dummy_f-1 -0.39146    1.20169  -0.326    0.745    
## pac_dummy_f1  -0.40231    0.64679  -0.622    0.534    
## ---
## Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1
## 
## Residual standard error: 2.935 on 1096 degrees of freedom
##   (74 observations deleted due to missingness)
## Multiple R-squared:  0.0004459,  Adjusted R-squared:  -0.001378 
## F-statistic: 0.2444 on 2 and 1096 DF,  p-value: 0.7832
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{summary}\NormalTok{(}\KeywordTok{lm}\NormalTok{(CCI }\OperatorTok{~}\StringTok{ }\NormalTok{ind2pty_dummy_f, }\DataTypeTok{data =}\NormalTok{ DiDsmall))}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## 
## Call:
## lm(formula = CCI ~ ind2pty_dummy_f, data = DiDsmall)
## 
## Residuals:
##     Min      1Q  Median      3Q     Max 
## -3.1870 -1.8367 -0.6887  1.0878 30.6592 
## 
## Coefficients:
##                   Estimate Std. Error t value Pr(>|t|)    
## (Intercept)        3.18704    0.08896  35.827   <2e-16 ***
## ind2pty_dummy_f-1 -3.18704    2.93559  -1.086    0.278    
## ind2pty_dummy_f1   0.32578    0.93214   0.349    0.727    
## ---
## Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1
## 
## Residual standard error: 2.934 on 1096 degrees of freedom
##   (74 observations deleted due to missingness)
## Multiple R-squared:  0.001187,   Adjusted R-squared:  -0.0006352 
## F-statistic: 0.6515 on 2 and 1096 DF,  p-value: 0.5215
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{summary}\NormalTok{(}\KeywordTok{lm}\NormalTok{(CRI }\OperatorTok{~}\StringTok{ }\NormalTok{ind_dummy_f, }\DataTypeTok{data =}\NormalTok{ DiDsmall))}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## 
## Call:
## lm(formula = CRI ~ ind_dummy_f, data = DiDsmall)
## 
## Residuals:
##      Min       1Q   Median       3Q      Max 
## -0.29085 -0.13371 -0.03591  0.09706  1.86780 
## 
## Coefficients:
##               Estimate Std. Error t value Pr(>|t|)    
## (Intercept)   0.290846   0.005898  49.317   <2e-16 ***
## ind_dummy_f-1 0.079369   0.079270   1.001    0.317    
## ind_dummy_f1  0.008152   0.048766   0.167    0.867    
## ---
## Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1
## 
## Residual standard error: 0.1936 on 1097 degrees of freedom
##   (73 observations deleted due to missingness)
## Multiple R-squared:  0.0009358,  Adjusted R-squared:  -0.0008857 
## F-statistic: 0.5138 on 2 and 1097 DF,  p-value: 0.5984
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{summary}\NormalTok{(}\KeywordTok{lm}\NormalTok{(CRI }\OperatorTok{~}\StringTok{ }\NormalTok{pac_dummy_f, }\DataTypeTok{data =}\NormalTok{ DiDsmall))}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## 
## Call:
## lm(formula = CRI ~ pac_dummy_f, data = DiDsmall)
## 
## Residuals:
##      Min       1Q   Median       3Q      Max 
## -0.29082 -0.13369 -0.03575  0.09653  1.86783 
## 
## Coefficients:
##               Estimate Std. Error t value Pr(>|t|)    
## (Intercept)   0.290823   0.005911  49.202   <2e-16 ***
## pac_dummy_f-1 0.086208   0.079265   1.088    0.277    
## pac_dummy_f1  0.005480   0.042662   0.128    0.898    
## ---
## Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1
## 
## Residual standard error: 0.1936 on 1097 degrees of freedom
##   (73 observations deleted due to missingness)
## Multiple R-squared:  0.00109,    Adjusted R-squared:  -0.0007316 
## F-statistic: 0.5983 on 2 and 1097 DF,  p-value: 0.5499
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{summary}\NormalTok{(}\KeywordTok{lm}\NormalTok{(CRI }\OperatorTok{~}\StringTok{ }\NormalTok{ind2pty_dummy_f, }\DataTypeTok{data =}\NormalTok{ DiDsmall))}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## 
## Call:
## lm(formula = CRI ~ ind2pty_dummy_f, data = DiDsmall)
## 
## Residuals:
##      Min       1Q   Median       3Q      Max 
## -0.29159 -0.13382 -0.03481  0.09631  1.86706 
## 
## Coefficients:
##                    Estimate Std. Error t value Pr(>|t|)    
## (Intercept)        0.291590   0.005870  49.674   <2e-16 ***
## ind2pty_dummy_f-1 -0.004768   0.193803  -0.025    0.980    
## ind2pty_dummy_f1  -0.020676   0.061538  -0.336    0.737    
## ---
## Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1
## 
## Residual standard error: 0.1937 on 1097 degrees of freedom
##   (73 observations deleted due to missingness)
## Multiple R-squared:  0.0001034,  Adjusted R-squared:  -0.00172 
## F-statistic: 0.05672 on 2 and 1097 DF,  p-value: 0.9449
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# No effects }

\KeywordTok{summary}\NormalTok{(}\KeywordTok{lm}\NormalTok{(CCI }\OperatorTok{~}\StringTok{ }\NormalTok{ind_dummy_f}\OperatorTok{*}\NormalTok{ind_d, }\DataTypeTok{data =}\NormalTok{ DiDsmall))}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## 
## Call:
## lm(formula = CCI ~ ind_dummy_f * ind_d, data = DiDsmall)
## 
## Residuals:
##     Min      1Q  Median      3Q     Max 
## -3.2595 -1.8493 -0.6567  1.0889 30.6486 
## 
## Coefficients:
##                       Estimate Std. Error t value Pr(>|t|)    
## (Intercept)          3.198e+00  8.969e-02  35.660   <2e-16 ***
## ind_dummy_f-1        2.759e+01  1.171e+02   0.236    0.814    
## ind_dummy_f1         4.038e+00  2.599e+01   0.155    0.877    
## ind_d                1.053e-05  1.088e-04   0.097    0.923    
## ind_dummy_f-1:ind_d -2.966e-04  1.187e-03  -0.250    0.803    
## ind_dummy_f1:ind_d   3.601e-05  2.896e-04   0.124    0.901    
## ---
## Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1
## 
## Residual standard error: 2.939 on 1093 degrees of freedom
##   (74 observations deleted due to missingness)
## Multiple R-squared:  0.0007875,  Adjusted R-squared:  -0.003783 
## F-statistic: 0.1723 on 5 and 1093 DF,  p-value: 0.9729
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{summary}\NormalTok{(}\KeywordTok{lm}\NormalTok{(CCI }\OperatorTok{~}\StringTok{ }\NormalTok{pac_dummy_f}\OperatorTok{*}\NormalTok{pac_d, }\DataTypeTok{data =}\NormalTok{ DiDsmall))}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## 
## Call:
## lm(formula = CCI ~ pac_dummy_f * pac_d, data = DiDsmall)
## 
## Residuals:
##     Min      1Q  Median      3Q     Max 
## -3.3006 -1.8479 -0.6341  1.0861 30.6491 
## 
## Coefficients:
##                       Estimate Std. Error t value Pr(>|t|)    
## (Intercept)          3.196e+00  8.987e-02  35.566   <2e-16 ***
## pac_dummy_f-1       -4.699e+00  4.269e+01  -0.110    0.912    
## pac_dummy_f1        -9.769e-01  2.786e+00  -0.351    0.726    
## pac_d               -8.187e-06  7.209e-05  -0.114    0.910    
## pac_dummy_f-1:pac_d  5.208e-05  4.407e-04   0.118    0.906    
## pac_dummy_f1:pac_d   1.622e-06  7.844e-05   0.021    0.984    
## ---
## Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1
## 
## Residual standard error: 2.939 on 1093 degrees of freedom
##   (74 observations deleted due to missingness)
## Multiple R-squared:  0.0005082,  Adjusted R-squared:  -0.004064 
## F-statistic: 0.1111 on 5 and 1093 DF,  p-value: 0.9899
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{summary}\NormalTok{(}\KeywordTok{lm}\NormalTok{(CCI }\OperatorTok{~}\StringTok{ }\NormalTok{ind2pty_dummy_f}\OperatorTok{*}\NormalTok{ind2pty_d, }\DataTypeTok{data =}\NormalTok{ DiDsmall))}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## 
## Call:
## lm(formula = CCI ~ ind2pty_dummy_f * ind2pty_d, data = DiDsmall)
## 
## Residuals:
##     Min      1Q  Median      3Q     Max 
## -3.2029 -1.8359 -0.6813  1.0865 30.6493 
## 
## Coefficients: (1 not defined because of singularities)
##                               Estimate Std. Error t value Pr(>|t|)    
## (Intercept)                  3.184e+00  8.904e-02  35.765   <2e-16 ***
## ind2pty_dummy_f-1           -1.944e+00  3.317e+00  -0.586    0.558    
## ind2pty_dummy_f1            -2.523e-01  1.133e+00  -0.223    0.824    
## ind2pty_d                   -1.313e-05  1.631e-05  -0.805    0.421    
## ind2pty_dummy_f-1:ind2pty_d         NA         NA      NA       NA    
## ind2pty_dummy_f1:ind2pty_d   6.468e-07  2.139e-05   0.030    0.976    
## ---
## Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1
## 
## Residual standard error: 2.935 on 1094 degrees of freedom
##   (74 observations deleted due to missingness)
## Multiple R-squared:  0.00252,    Adjusted R-squared:  -0.001127 
## F-statistic: 0.691 on 4 and 1094 DF,  p-value: 0.5983
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{summary}\NormalTok{(}\KeywordTok{lm}\NormalTok{(CRI }\OperatorTok{~}\StringTok{ }\NormalTok{ind_dummy_f}\OperatorTok{*}\NormalTok{ind_d, }\DataTypeTok{data =}\NormalTok{ DiDsmall))}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## 
## Call:
## lm(formula = CRI ~ ind_dummy_f * ind_d, data = DiDsmall)
## 
## Residuals:
##      Min       1Q   Median       3Q      Max 
## -0.29128 -0.13350 -0.03540  0.09671  1.86737 
## 
## Coefficients:
##                       Estimate Std. Error t value Pr(>|t|)    
## (Intercept)          2.913e-01  5.908e-03  49.301   <2e-16 ***
## ind_dummy_f-1       -3.567e+00  7.717e+00  -0.462    0.644    
## ind_dummy_f1         1.324e+00  1.713e+00   0.773    0.440    
## ind_d                9.312e-06  7.168e-06   1.299    0.194    
## ind_dummy_f-1:ind_d  2.750e-05  7.826e-05   0.351    0.725    
## ind_dummy_f1:ind_d   4.280e-06  1.908e-05   0.224    0.823    
## ---
## Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1
## 
## Residual standard error: 0.1937 on 1094 degrees of freedom
##   (73 observations deleted due to missingness)
## Multiple R-squared:  0.003215,   Adjusted R-squared:  -0.001341 
## F-statistic: 0.7057 on 5 and 1094 DF,  p-value: 0.6192
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{summary}\NormalTok{(}\KeywordTok{lm}\NormalTok{(CRI }\OperatorTok{~}\StringTok{ }\NormalTok{pac_dummy_f}\OperatorTok{*}\NormalTok{pac_d, }\DataTypeTok{data =}\NormalTok{ DiDsmall))}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## 
## Call:
## lm(formula = CRI ~ pac_dummy_f * pac_d, data = DiDsmall)
## 
## Residuals:
##      Min       1Q   Median       3Q      Max 
## -0.29115 -0.13375 -0.03518  0.09523  1.86789 
## 
## Coefficients:
##                       Estimate Std. Error t value Pr(>|t|)    
## (Intercept)          2.908e-01  5.923e-03  49.089   <2e-16 ***
## pac_dummy_f-1       -1.586e+00  2.815e+00  -0.563    0.573    
## pac_dummy_f1         1.285e-01  1.837e-01   0.700    0.484    
## pac_d               -1.018e-06  4.753e-06  -0.214    0.830    
## pac_dummy_f-1:pac_d  1.806e-05  2.906e-05   0.621    0.534    
## pac_dummy_f1:pac_d   2.422e-06  5.172e-06   0.468    0.640    
## ---
## Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1
## 
## Residual standard error: 0.1938 on 1094 degrees of freedom
##   (73 observations deleted due to missingness)
## Multiple R-squared:  0.001886,   Adjusted R-squared:  -0.002676 
## F-statistic: 0.4135 on 5 and 1094 DF,  p-value: 0.8396
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{summary}\NormalTok{(}\KeywordTok{lm}\NormalTok{(CRI }\OperatorTok{~}\StringTok{ }\NormalTok{ind2pty_dummy_f}\OperatorTok{*}\NormalTok{ind2pty_d, }\DataTypeTok{data =}\NormalTok{ DiDsmall))}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## 
## Call:
## lm(formula = CRI ~ ind2pty_dummy_f * ind2pty_d, data = DiDsmall)
## 
## Residuals:
##      Min       1Q   Median       3Q      Max 
## -0.29198 -0.13371 -0.03515  0.09587  1.86716 
## 
## Coefficients: (1 not defined because of singularities)
##                               Estimate Std. Error t value Pr(>|t|)    
## (Intercept)                  2.915e-01  5.877e-03  49.595   <2e-16 ***
## ind2pty_dummy_f-1            4.441e-02  2.191e-01   0.203    0.839    
## ind2pty_dummy_f1             1.013e-02  7.483e-02   0.135    0.892    
## ind2pty_d                   -5.197e-07  1.077e-06  -0.483    0.630    
## ind2pty_dummy_f-1:ind2pty_d         NA         NA      NA       NA    
## ind2pty_dummy_f1:ind2pty_d   1.180e-06  1.413e-06   0.835    0.404    
## ---
## Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1
## 
## Residual standard error: 0.1938 on 1095 degrees of freedom
##   (73 observations deleted due to missingness)
## Multiple R-squared:  0.0007914,  Adjusted R-squared:  -0.002859 
## F-statistic: 0.2168 on 4 and 1095 DF,  p-value: 0.9291
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# No effects }

\CommentTok{# Next: Lag, write up. }
\end{Highlighting}
\end{Shaded}

\end{document}