### 2020-04-03 draft-master/inst/rmarkdown/templates/jop/skeleton/skeleton.tex

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}

\documentclass{jopsubmission}

%% Math support
\usepackage{amssymb} % Note: loads amsfonts
\usepackage{amsmath}

\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{textcomp} % provide euro and other symbols

% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}

\usepackage{xcolor}

\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available

\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={Your Title},
          hidelinks,
  pdfcreator={LaTeX via pandoc}}

\urlstyle{same} % disable monospaced font for URLs

%% For tables
\usepackage{booktabs}

%% For figures
\usepackage{graphicx}
\usepackage{grffile}

\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

% We don't want a date in there
\date{\vspace{-5ex}}

    \title{Your Title}

% We need to remove section numbering
\setcounter{secnumdepth}{-\maxdimen}

\begin{document}

    \maketitle
            \begin{abstract}
        Your abstract should go here. It should not exceed 150 words.
        \end{abstract}

Your introduction should go at the very beginning. It should not have a
section heading. Note that your initial submission should be anonymized.
This is, in part (I cannot stop you from referencing your identity in
the text), enforced by the YAML option ``anonymized'', which is true by
default. If you set it to false, your name (as provided in the
``author'' option) will be displayed via
\texttt{\textbackslash{}maketitle}.

Short article submissions must be 10 pages or less, including tables,
figures, footnotes, and references, but excluding the title page,
abstract page, and online appendix. Research articles must be 35 pages
or less (same counting).

\section{Section headings should be
short}\label{section-headings-should-be-short}

Section headings should not take up more than a single line of text.
There should be no section heading for the introduction. Do not use
headings immediately after one another without text in between. You
generally should not use headings at the sub-sub-heading level or below.
Do not refer to sections by number, as sections are not numbered.

\end{document}